---
title: "BIOL312_FinalProj"
author: "Minh Tran"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---
### Source: Chase, M.A and Dummer, G.M. (1992), "The Role of Sports as a Social Determinant for Children," Research Quarterly for Exercise and Sport, 63, 418-424. 


## Relationship between count data of goals in different grades


### Load up packages and dataset
```{r, message=FALSE, warning=FALSE}
library('tidyverse')
data <- read_csv('youth_social_preferences.csv')
```

### Inspect dataset
```{r}
glimpse(data) #checking out data types and observations
summary(data) #checking to see if there are any NAs
```

### A. Statistical hypotheses

`Alpha: 0.05`

`H0: There is no significant relationship between the count of goals and grades. `

`H1: There is a significant relationship between the count of goals and grades. ` 

### B. Cross tabulation of count of goals in different grades
```{r}
xt1 <- xtabs(~Grade + Goals, data = data) #cross tab
xt1 
prop.table(xt1, margin = 1) #proportion table
```

### C. Expected chi-square values and running chi-square
```{r}
chisq.test(xt1)$expected #expected chi square values
chisq.test(xt1, simulate.p.value = TRUE) #chi square test with simulated p-value
```

### D. Interpretation

`The p-value is higher than the alpha of 0.05. Not significant -> Fail to reject the null hypothesis and conclude that there is no significant relationship between the count of goals and grades.`

### E. Barplot visuals
```{r}
#getting counts and plotting a barplot of counts of each Grade and Goal combination
data %>% 
  count(Grade, Goals) %>% 
  ggplot(aes(x = Grade, y = n, fill = Goals)) +
  geom_bar(position = "dodge",stat = "identity") +
  theme_bw() +
  labs(title = 'Count data of Goals vs. Grades')+
  xlab('Grades') +
  ylab('Count')
```

```{r}
#getting counts and plotting a stacked barplot of percentages of each Grade and Goal combination
data %>% 
  count(Grade, Goals) %>% 
  ggplot(aes(x = Grade, y = n, fill = Goals)) +
  geom_bar(position = "fill",stat = "identity") +
  theme_bw() +
  xlab('Grades') +
  ylab('Count (%)') +
  labs(title = 'Count percentage of Goals vs. Grades') +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip()
```

---

## Relationship between count data of goals in different communities

### A. Statistical hypotheses

`Alpha: 0.05`

`H0: There is no significant relationship between the count of goals and communities. ` 

`H1: There is a significant relationship between the goals and communities.` 

### B. Cross tabulation of count of goals in different communities
```{r}
xt2 <- xtabs(~Urban.Rural + Goals, data = data) #cross tab
xt2
prop.table(xt2, margin = 1) #table of proportions
```

### C. Expected chi-square values and running chi-square
```{r}
chisq.test(xt2)$expected #expected chi-square values
chisq.test(xt2, simulate.p.value = TRUE) #chi-square test with simulated p-value
```

### D. Interpretation

`The p-value is lower than the alpha of 0.05. Significant -> Reject the null hypothesis and conclude that there is a significant relationship between the count of goals and communities. `

### E. Barplot visuals

```{r}
#getting counts and plotting a barplot of counts of each Community and Goal combination
data %>% 
  count(Urban.Rural, Goals) %>% 
  ggplot(aes(x = Urban.Rural, y = n, fill = Goals)) +
  geom_bar(position = "dodge",stat = "identity") +
  theme_bw() +
  labs(title = 'Count data of Goals vs. Community') +
  xlab('Community') +
  ylab('Count') 
```

```{r}
#getting counts and plotting a stacked barplot of percentages of each Community and Goal combination
data %>% 
  count(Urban.Rural, Goals) %>% 
  ggplot(aes(x = Urban.Rural, y = n, fill = Goals)) +
  geom_bar(position = "fill",stat = "identity") +
  theme_bw() +
  xlab('Community') +
  ylab('Count (%)') + 
  labs(title = 'Count percentage of Goals vs. Community') +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() 
```
